\begin{filecontents*}{\jobname.xmpdata}
  \Title{Building a Ray-Traced Rendering Engine on Sparse Voxel Grids}
  \Author{10834225}
  \Language{en-GB}
  \Copyrighted{True}
\end{filecontents*}

\documentclass{extra}

%%%%%%%%%%%%%%%%%% PACKAGES AND COMMANDS %%%%%%%%%%%%%%%%%%
\usepackage{graphicx,psfrag,color} % for postscript graphics files
  \graphicspath{ {./figures/} }
\usepackage{amsmath}               % assumes amsmath package installed
  \allowdisplaybreaks[1]           % allow eqnarrays to break across pages
\usepackage{amssymb}               % assumes amsmath package installed
\usepackage{url}                   % format hyperlinks correctly
\usepackage{rotating}              % allow portrait figures and tables
\usepackage{multirow}              % allows merging of rows in tables
\usepackage{lscape}                % allows pages to be typeset in landscape mode
\usepackage{tabularx}              % allows fixed width tables
\usepackage{verbatim}              % enhanced version of built-in verbatim environment
\usepackage{footnote}              % allows more control over footnote environments
\usepackage{float}                 % allows H option on floats to force here placement
\usepackage{booktabs}              % improve table line spacing
\usepackage{lipsum}                % for adding dummy text here
\usepackage[base]{babel}           % for proper hypthenation in lipsum sections
\usepackage{subcaption}            % for multiple sub-figures in a single float
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{color}
\usepackage[inkscapelatex=false]{svg}
\usepackage[acronym]{glossaries}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{multicol}
\usepackage{bm}
\usepackage{enumitem}
\newtheorem{theorem}{Theorem}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\sus}[1]{$^{\mbox{\scriptsize #1}}$} % superscript in text (e.g. 1st)
\newcommand{\sub}[1]{$_{\mbox{\scriptsize #1}}$} % subscript in text
\AtBeginEnvironment{quote}{\par\singlespacing\small}

%%%%%%%%%%%%%%%%%% REFERENCES SETUP %%%%%%%%%%%%%%%%%%
\usepackage[style=ieee,backend=biber,backref=true,hyperref=auto,backend=bibtex]{biblatex}
% \DefineBibliographyStrings{english}{backrefpage = {cited on p\adddot},  backrefpages = {cited on pp\adddot}}
\addbibresource{ref.bib}

\makeglossary
\newacronym{gcd}{GCD}{Greatest Common Divisor}

%%%%%%%%%%%%%%%%%% START DOCUMENT %%%%%%%%%%%%%%%%%%
\begin{document}
\makeatletter
\title{\xmp@Title}
\studentid{\xmp@Author}
\makeatother

\course{Computer Science}
\submitdate{2024}
\wordcount{many}
\maketitle



%%%%%%%%%%%%%%%%%% LISTS OF CONTENT %%%%%%%%%%%%%%%%%%
\uomtoc
% other lists are not required, but can include \uomlof and \uomlot if really want to

\uomlof

\uomlot

%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%
\begin{abstract} % put abstract here. Limit is 1 page.
Rendering engines are critical in computer graphics, serving as the backbone of visualisation for digital media, including video games, simulations, virtual reality, and animated films. These engines are responsible for converting 3D models, textures, and lighting information into the compelling images users see on their screens. A rendering engine's efficiency and capabilities directly influence the digital experience's realism and interactivity.

This thesis explores the development and implementation of a ray-traced voxel rendering engine utilising sparse grids to enhance real-time rendering capabilities. The focus is on leveraging sparse data structures to manage volumetric data efficiently, thus allowing for intricate rendering details and high performance. The research investigates various techniques in ray tracing to optimise the rendering process. The outcomes demonstrate that sparse voxel grids combined with discrete signed distance fields can significantly reduce memory usage while maintaining rendering speed. This work contributes to computer graphics and game development by providing insights into the application of sparse data structures in real-time rendering environments.
\end{abstract}%
\clearpage


%%%%%%%%%%%%%%%%%% DECLARATIONS %%%%%%%%%%%%%%%%%%
\uomdeclarations % Don't need unless final thesis


\begin{uomacknowledgements}
  I would like to express my gratitude to my supervisor, Prof. Steve Pettifer, for his guidance, encouragement, and expertise throughout this research project. My thanks also go to my partner Diana, whose support and understanding have been my anchor in my academic pursuits. I also extend my deepest appreciation to my parents, whose love and belief in my potential have been my constant motivation.
\end{uomacknowledgements}

%%%%%%%%%%%%%%%%%% CONTENT %%%%%%%%%%%%%%%%%%
\include{chapters/intro.tex}
\include{chapters/background.tex}
\include{chapters/methodology.tex}
\include{chapters/results.tex}
\include{chapters/conclusion.tex}

%%%%%%%%%%%%%%%%%% REFERRENCES %%%%%%%%%%%%%%%%%%
\printbibliography[title={References},heading=bibintoc]


%%%%%%%%%%%%%%%%%% ACRONYMS %%%%%%%%%%%%%%%%%%
\newpage
\printglossary[type=\acronymtype]

\end{document}
